---
title: "Analysis"
author: "Matthew Perrotta"
date: "April 20, 2019"
output: html_document
---

Data analysis for MPH Epidemiology thesis requirement.

Research Question: What is the best predictor of standardized malaria rate (SMR) in the Malaysian regions of Sarawak and Sabah.

Methods:
-Exploratory data analysis and visualizations
-multiple linear regression
-more advanced linear regression methods
-GAM

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Libraries
```{r echo = T, message = FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(mgcv)
library(corrplot)
library(Hmisc)
library(caret)
library(PerformanceAnalytics)
library(car)
library(olsrr)
#library(MASS)
```

READ: https://www.nature.com/articles/s41598-017-02560-z

### Import and clean data
```{r}
swk = read.csv('./data/data_full_sar.csv') %>% 
  group_by(district, year) %>% 
  mutate(district_area = 
           CropRainfed.total.area + 
           Herbaceous.total.area + 
           TreeShrub.total.area +
           MosCrop.total.area +
           MosNatural.total.area + 
           BroadEvgClop.total.area +
           Shrub.total.area +
           ShrubEvg.total.area +
           SparseVeg.total.area +
           FloodFresh.total.area +
           FloodSalt.total.area +
           Urban.total.area +
           Water.total.area +
           CropIrrigate.total.area 
         ) %>% 
  mutate(pop_density = population/district_area) %>% 
  ungroup() %>% 
  group_by(district) %>%
  mutate(CropRainfed.prop = CropRainfed.total.area/district_area,
         Herbaceous.prop = Herbaceous.total.area/district_area,
         TreeShrub.prop = TreeShrub.total.area/district_area,
         MosCrop.prop = MosCrop.total.area/district_area,
         MosNatural.prop = MosNatural.total.area/district_area,
         BroadEvgClop.prop = BroadEvgClop.total.area/district_area,
         Shrub.prop = Shrub.total.area/district_area,
         ShrubEvg.prop = ShrubEvg.total.area/district_area,
         SparseVeg.prop = SparseVeg.total.area/district_area,
         FloodFresh.prop = FloodFresh.total.area/district_area,
         FloodSalt.prop = FloodSalt.total.area/district_area,
         Urban.prop = Urban.total.area/district_area,
         Water.prop = Water.total.area/district_area,
         CropIrrigate.prop = CropIrrigate.total.area/district_area) %>% 
  janitor::clean_names() %>%
  ungroup() %>%  
  select(-c('x',
            'population_year',
            'cases_year',
            'expected',
            'sd')) %>% 
  rename('smr' = 'sir',
         'prec_mean' = 'mean',
         'cases' = 'case_number') %>% 
  mutate(year = as.factor(year)) %>% 
  na.omit()


sbh = read.csv('./data/dataFull.csv') %>% 
  group_by(District, Year) %>% 
  mutate(district_area = 
           CropRainfed.total.area + 
           Herbaceous.total.area + 
           TreeShrub.total.area +
           MosCrop.total.area +
           MosNatural.total.area + 
           BroadEvgClop.total.area +
           Shrub.total.area +
           ShrubEvg.total.area +
           SparseVeg.total.area +
           FloodFresh.total.area +
           FloodSalt.total.area +
           Urban.total.area +
           Water.total.area +
           CropIrrigate.total.area +
           Grass.total.area +
           MosTreeHerb.total.area
         ) %>%
  mutate(pop_density = population/district_area) %>% 
  ungroup() %>% 
  group_by(District) %>%
  mutate(CropRainfed.prop = CropRainfed.total.area/district_area,
         Herbaceous.prop = Herbaceous.total.area/district_area,
         TreeShrub.prop = TreeShrub.total.area/district_area,
         MosCrop.prop = MosCrop.total.area/district_area,
         MosNatural.prop = MosNatural.total.area/district_area,
         BroadEvgClop.prop = BroadEvgClop.total.area/district_area,
         Shrub.prop = Shrub.total.area/district_area,
         ShrubEvg.prop = ShrubEvg.total.area/district_area,
         SparseVeg.prop = SparseVeg.total.area/district_area,
         FloodFresh.prop = FloodFresh.total.area/district_area,
         FloodSalt.prop = FloodSalt.total.area/district_area,
         Urban.prop = Urban.total.area/district_area,
         Water.prop = Water.total.area/district_area,
         CropIrrigate.prop = CropIrrigate.total.area/district_area,
         Grass.prop = Grass.total.area/district_area,
         MosTreeHerb.prop = MosTreeHerb.total.area/district_area) %>% 
  janitor::clean_names() %>%
  ungroup() %>%
  select(-c('x', 
            'disease', 
            'number_deaths', 
            'mortality_rates',
            'prevalence',
            'incidence_rate')) %>% 
  rename('cases' = 'number_cases') %>% 
  mutate(year = as.factor(year)) %>% 
  na.omit()
```
Observations with `NA` values were omitted from the data set

Land Cover variable discriptions:
`crop_rainfed` - Cropland, rainfed
`crop_irrigate` - Cropland, irrigated or post-flooding
`mos_crop` - Mosaic cropland (>50%) / natural vegetation (tree, shrub, herbaceous cover)(<50%) 
`mos_natural` - Mosaic natural vegetation (tree, shrub, herbaceous cover) (>50%) / cropland (<50%) 
`mos_tree_herb` - Mosaic tree and shrub (>50%) / herbaceous cover (<50%)
`herbaceous` - Herbaceous cover
`tree_shrub` - Tree or shrub cover	
`broad_evg_clop` - Tree cover, broadleaved, evergreen, closed to open (>15%)
`shrub` - Shrubland
`shrub_evg` - Shrubland evergreen	
`sparse_veg` - Sparse vegetation (tree, shrub, herbaceous cover) (<15%)	
`flood_fresh` - Tree cover, flooded, fresh or brakish water	
`flood_salt` - Tree cover, flooded, saline water	
`grass` - Grassland
`urban` - Urban areas	
`water` - Water bodies	


The Sarawak dataset does not have the following variables: 
`grass_n_patches`
`grass_patch_density`
`grass_total_area`
`grass_prop`
`mos_tree_herb_n_patches`
`mos_tree_herb_patch_density`
`mos_tree_herb_total_area`
`mos_tree_herb_prop`

```{r}
summary(sbh)
```


Correlation 
```{r}
sbh_sub = sbh %>% 
  select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "shrub_evg_prop", "shrub_evg_n_patches", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density","crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "shrub_evg_patch_density", "prec_mean"))

res_sub <- rcorr(as.matrix(sbh_sub))

corrplot(res_sub$r, type = "lower", 
        p.mat = res_sub$P, sig.level = 0.01, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex = 0.5)

swk_sub = swk %>%
  select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "shrub_evg_prop", "shrub_evg_n_patches", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density","crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "shrub_evg_patch_density", "prec_mean"))

res_sub2 <- rcorr(as.matrix(swk_sub))

corrplot(res_sub2$r, type = "lower", 
        p.mat = res_sub2$P, sig.level = 0.01, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex=0.5)
```

Correlations between land use total area variables
```{r}
sbh_lu = sbh %>% 
  select(c("broad_evg_clop_total_area", 
           "crop_rainfed_total_area", 
           "herbaceous_total_area", 
           "shrub_total_area",
           "shrub_evg_total_area",
           "urban_total_area", 
           "tree_shrub_total_area", 
           "mos_crop_total_area", 
           "mos_natural_total_area", 
           "sparse_veg_total_area", 
           "flood_fresh_total_area", 
           "flood_salt_total_area", 
           "water_total_area", 
           "crop_irrigate_total_area",
           "grass_total_area",
           "mos_tree_herb_total_area"))

res_lu <- rcorr(as.matrix(sbh_lu))

corrplot(res_lu$r, type = "lower", 
        p.mat = res_lu$P, sig.level = 0.01, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex = 0.5)

swk_lu = swk %>% 
  select(c("broad_evg_clop_total_area", 
           "crop_rainfed_total_area", 
           "herbaceous_total_area", 
           "shrub_total_area",
           "shrub_evg_total_area",
           "urban_total_area", 
           "tree_shrub_total_area", 
           "mos_crop_total_area", 
           "mos_natural_total_area", 
           "sparse_veg_total_area", 
           "flood_fresh_total_area", 
           "flood_salt_total_area", 
           "water_total_area", 
           "crop_irrigate_total_area"))

res_lu2 <- rcorr(as.matrix(swk_lu))

corrplot(res_lu2$r, type = "lower", 
        p.mat = res_lu2$P, sig.level = 0.01, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex=0.5)
```


```{r}
sbh.x = model.matrix(smr~., sbh_sub, na.action = NULL)[,-1]
sbh.y = sbh_sub$smr

swk.x = model.matrix(smr~., swk_sub)[,-1]
swk.y = swk_sub$smr
```


Scatterplot of predictors vs. outcome
```{r}
featurePlot(sbh.x, 
            sbh.y, 
            plot = "scatter", 
            labels = c("Predictors","SMR"),
            type = c('p'))

featurePlot(swk.x, 
            swk.y, 
            plot = "scatter", 
            labels = c("Predictors","SMR"),
            type = c('p'))
```


### Checking for Multicolliearity with Tolerance and VIF

```{r}
ols_vif_tol(lm(smr ~ broad_evg_clop_total_area + 
           crop_rainfed_total_area + 
           herbaceous_total_area + 
           shrub_total_area +
           shrub_evg_total_area +
           urban_total_area + 
           tree_shrub_total_area + 
           mos_crop_total_area + 
           mos_natural_total_area + 
           sparse_veg_total_area + 
           flood_fresh_total_area + 
           flood_salt_total_area + 
           water_total_area + 
           crop_irrigate_total_area +
           mos_tree_herb_total_area +
           grass_total_area, 
           data = sbh))

ols_vif_tol(lm(smr ~ broad_evg_clop_total_area + 
           crop_rainfed_total_area + 
           herbaceous_total_area + 
           shrub_total_area +
           shrub_evg_total_area +
           urban_total_area + 
           tree_shrub_total_area + 
           mos_crop_total_area + 
           mos_natural_total_area + 
           sparse_veg_total_area + 
           flood_fresh_total_area + 
           flood_salt_total_area + 
           water_total_area + 
           crop_irrigate_total_area, 
           data = swk))
```

I will create general cropland variable to decrease the number of total variables and account for multicollinearity between many land use variables. This is based on the article "Comparison of Global Land Cover Datasets for
Cropland Monitoring".

```{r}
sbh = sbh %>% 
  group_by(district, year) %>% 
  mutate(agri_total_area = crop_rainfed_total_area + 
           crop_irrigate_total_area +
           herbaceous_total_area +
           tree_shrub_total_area +
           mos_crop_total_area + 
           mos_natural_total_area) %>% 
  mutate(agri_prop = agri_total_area/district_area)

swk = swk %>% 
  group_by(district, year) %>% 
  mutate(agri_total_area = crop_rainfed_total_area + 
           crop_irrigate_total_area +
           herbaceous_total_area +
           tree_shrub_total_area +
           mos_crop_total_area + 
           mos_natural_total_area) %>% 
  mutate(agri_prop = agri_total_area/district_area)
```

Rechecking VIF after new variable for agriculture is created
```{r}
ols_vif_tol(lm(smr ~ broad_evg_clop_total_area + 
           shrub_total_area +
           shrub_evg_total_area +
           urban_total_area + 
           agri_total_area + 
           sparse_veg_total_area + 
           flood_fresh_total_area + 
           flood_salt_total_area + 
           water_total_area + 
           mos_tree_herb_total_area +
           grass_total_area, 
           data = sbh))

ols_vif_tol(lm(smr ~ broad_evg_clop_total_area + 
           shrub_total_area +
           shrub_evg_total_area +
           urban_total_area + 
           agri_total_area + 
           sparse_veg_total_area + 
           flood_fresh_total_area + 
           flood_salt_total_area + 
           water_total_area, 
           data = swk))
```


### EDA and Visualization

Looking at the distribution of the outcome (SMR) for both Sarawak and Sabah
```{r warning=FALSE}
ggplot(data = swk, aes(x = smr)) +  
  geom_histogram() +
  labs(title = 'SMR distribution, Sarawak')

ggplot(data = sbh, aes(x = smr)) +  
  geom_histogram() +
  labs(title = 'SMR distribution, Sabah') 
```
Distribution seems to follow a poisson distribution. When running GAM, incorporate Poisson regression. Also, SMR by its nature of being a rate follows  poisson regression.

### Fitting Poisson Regression Model for Sabah and Sarawak

Sabah
```{r}
sbh.glm = glm(smr ~ 
                broad_evg_clop_n_patches +
                broad_evg_clop_patch_density +
                broad_evg_clop_prop +
                crop_irrigate_n_patches +
                crop_irrigate_patch_density +
                crop_rainfed_n_patches + 
                crop_rainfed_patch_density +
                flood_fresh_n_patches +
                flood_fresh_patch_density +
                flood_fresh_prop +
                flood_salt_n_patches +
                flood_salt_patch_density +
                flood_salt_prop +
                herbaceous_n_patches +
                herbaceous_patch_density +
                mos_crop_n_patches +
                mos_crop_patch_density +
                mos_natural_n_patches +
                mos_natural_patch_density +
                shrub_n_patches +
                shrub_patch_density +
                shrub_prop +
                shrub_evg_n_patches +
                shrub_evg_patch_density +
                shrub_evg_prop +
                sparse_veg_n_patches +
                sparse_veg_patch_density +
                sparse_veg_prop +
                tree_shrub_n_patches +
                tree_shrub_patch_density + 
                grass_n_patches +
                grass_patch_density +
                grass_prop +
                urban_n_patches +
                urban_patch_density +
                urban_prop +
                water_n_patches +
                water_patch_density +
                agri_prop +
                water_prop +
                pop_density +
                prec_mean +
                year +
                district, 
              data = sbh, family = poisson(link = log)
              )

summary(sbh.glm)
plot(sbh.glm)
```


Checking variance and mean for both datasets
```{r}
labels = c('Region', 'Mean', 'Variance')
Sabah = c(mean(sbh$smr, na.rm = T), var(sbh$smr, na.rm = T))
Sarawak = c(mean(swk$smr, na.rm = T), var(swk$smr, na.rm = T))

mean_and_variance = rbind(Sabah, Sarawak)
```
For both regions the mean and variance differ, indicating the need to use a negative binomal distribution instead of poisson.


### Fitting GAM for Sabah and Sarawak

Poisson models will be used as the outcome is SMR

GAM model, Land Cover Proportions
```{r}
fit.prop = bam(smr ~ s(broad_evg_clop_prop) +
                  s(agri_prop, k = 5) +
                  s(mos_tree_herb_prop, k = 5) +
                  s(grass_prop, k = 3) +
                  s(sparse_veg_prop, k = 3) +
                  s(shrub_prop) +
                  s(shrub_evg_prop, k = 3) +
                  s(flood_fresh_prop) +
                  s(flood_salt_prop) +
                  s(urban_prop) +
                  s(water_prop) +
                  year +
                  district,
                data = sbh,
                family = neg.bin(1)
               )

summary(fit.prop)
gam.check(fit.prop)
plot(fit.prop)
```

```{r}
fit.prop.int = bam(smr ~ s(broad_evg_clop_prop) +
                  s(agri_prop, k = 5) +
                  s(mos_tree_herb_prop, k = 5) +
                  s(grass_prop, k = 5) +
                  s(sparse_veg_prop, k = 2) +
                  s(shrub_prop) +
                  s(shrub_evg_prop, k = 3) +
                  s(flood_fresh_prop) +
                  s(flood_salt_prop) +
                  s(urban_prop) +
                  s(water_prop) +
                  #ti(broad_evg_clop_prop, by = district) +
                  #ti(agri_prop, district) +
                  #ti(mos_tree_herb_prop, district) +
                  #ti(grass_prop, district) +
                  #ti(sparse_veg_prop, district) +
                  #ti(shrub_prop, district) +
                  #ti(shrub_evg_prop, district) +
                  #ti(flood_fresh_prop, district) +
                  #ti(flood_salt_prop, district) +
                  #ti(urban_prop, district) +
                  #ti(water_prop, district) +
                  year +
                  district,
                data = sbh,
                family = poisson(link = log)
               )

summary(fit.prop.int)
gam.check(fit.prop.int)
plot(fit.prop.int)
```


GAM, with Poisson

Model takes too long to run, so I will be removing all `_n_patches` variables. These variables are accounted for 
```{r}
sbh.fit2 = bam(smr ~ 
                s(broad_evg_clop_n_patches, k = 5) +
                s(broad_evg_clop_patch_density, k = 5) +
                s(broad_evg_clop_prop, k = 5) +
                s(crop_irrigate_n_patches, k = 5) +
                s(crop_irrigate_patch_density, k = 5) +
                s(crop_irrigate_prop, k = 5) +
                s(crop_rainfed_n_patches, k = 5) + 
                s(crop_rainfed_patch_density, k = 5) +
                s(crop_rainfed_prop, k = 5) +
                s(flood_fresh_n_patches, k = 5) +
                s(flood_fresh_patch_density, k = 5) +
                s(flood_fresh_prop, k = 5) +
                s(flood_salt_n_patches, k = 5) +
                s(flood_salt_patch_density, k = 5) +
                s(flood_salt_prop, k = 5) +
                s(herbaceous_n_patches, k = 5) +
                s(herbaceous_patch_density, k = 5) +
                s(herbaceous_prop, k = 5) +
                s(mos_crop_n_patches, k = 5) +
                s(mos_crop_patch_density, k = 5) +
                s(mos_crop_prop, k = 5) +
                s(mos_natural_n_patches, k = 5) +
                s(mos_natural_patch_density, k = 5) +
                s(mos_natural_prop, k = 5) +
                s(shrub_n_patches, k = 5) +
                s(shrub_patch_density, k = 5) +
                s(shrub_prop, k = 5) +
                s(shrub_evg_n_patches, k = 5) +
                s(shrub_evg_patch_density, k = 5) +
                s(shrub_evg_prop, k = 5) +
                s(sparse_veg_n_patches, k = 5) +
                s(sparse_veg_patch_density, k = 5) +
                s(sparse_veg_prop, k = 5) +
                s(tree_shrub_n_patches, k = 5) +
                s(tree_shrub_patch_density, k = 5) + 
                s(tree_shrub_prop, k = 5) +
                s(grass_n_patches, k =5) +
                s(grass_patch_density, k = 5) +
                s(grass_prop, k = 5) +
                s(urban_n_patches, k = 5) +
                s(urban_patch_density, k = 5) +
                s(urban_prop, k = 5) +
                s(water_n_patches, k = 5) +
                s(water_patch_density, k = 5) +
                s(water_prop, k = 5) +
                s(prec_mean, k = 5) +
                year +
                district, 
              data = sbh, method = 'REML',
              family = poisson(link = log)
              ) 

summary(sbh.fit2)
gam.check(sbh.fit2)
```

Inspect model:
gam.check()
concurvity()
VIF/tolerance
summary()
coef()
```{r}
summary(sbh_fit)
gam.check(gam_sbh)
```


