---
title: "Analysis"
author: "Matthew Perrotta"
date: "April 20, 2019"
output: html_document
---

Data analysis for MPH Epidemiology thesis requirement.

Research Question: What is the best predictor of standardized malaria rate (SMR) in the Malaysian regions of Sarawak and Sabah.

Methods:
-Exploratory data analysis and visualizations
-multiple linear regression
-more advanced linear regression methods
-GAM

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Libraries
```{r echo = T, message = FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(mgcv)
library(corrplot)
library(Hmisc)
library(caret)
library(PerformanceAnalytics)
library(car)
library(olsrr)
#library(MASS)
#library(raster)
```

READ: https://www.nature.com/articles/s41598-017-02560-z

### Import and clean data

Two datasets are created, `swk` for Sarawak and `sbh` for Sabah. Variables that are not needed are removed.
```{r}
swk = read.csv('./data/data_full_sar.csv') %>% 
  group_by(district, year) %>% 
  mutate(district_area = 
           CropRainfed.total.area + 
           Herbaceous.total.area + 
           TreeShrub.total.area +
           MosCrop.total.area +
           MosNatural.total.area + 
           BroadEvgClop.total.area +
           Shrub.total.area +
           ShrubEvg.total.area +
           SparseVeg.total.area +
           FloodFresh.total.area +
           FloodSalt.total.area +
           Urban.total.area +
           Water.total.area +
           CropIrrigate.total.area 
         ) %>% 
  mutate(pop_density = population/district_area,
         log_smr = log(SIR)) %>% 
  ungroup() %>% 
  group_by(district) %>%
  mutate(CropRainfed.prop = CropRainfed.total.area/district_area,
         Herbaceous.prop = Herbaceous.total.area/district_area,
         TreeShrub.prop = TreeShrub.total.area/district_area,
         MosCrop.prop = MosCrop.total.area/district_area,
         MosNatural.prop = MosNatural.total.area/district_area,
         BroadEvgClop.prop = BroadEvgClop.total.area/district_area,
         Shrub.prop = Shrub.total.area/district_area,
         ShrubEvg.prop = ShrubEvg.total.area/district_area,
         SparseVeg.prop = SparseVeg.total.area/district_area,
         FloodFresh.prop = FloodFresh.total.area/district_area,
         FloodSalt.prop = FloodSalt.total.area/district_area,
         Urban.prop = Urban.total.area/district_area,
         Water.prop = Water.total.area/district_area,
         CropIrrigate.prop = CropIrrigate.total.area/district_area) %>% 
  janitor::clean_names() %>%
  ungroup() %>%  
  select(-c('x',
            'population_year',
            'cases_year',
            'expected',
            'sd')) %>% 
  rename('smr' = 'sir',
         'prec_mean' = 'mean',
         'cases' = 'case_number') %>% 
  na.omit()


sbh = read.csv('./data/dataFull.csv') %>% 
  group_by(District, Year) %>% 
  mutate(district_area = 
           CropRainfed.total.area + 
           Herbaceous.total.area + 
           TreeShrub.total.area +
           MosCrop.total.area +
           MosNatural.total.area + 
           BroadEvgClop.total.area +
           Shrub.total.area +
           ShrubEvg.total.area +
           SparseVeg.total.area +
           FloodFresh.total.area +
           FloodSalt.total.area +
           Urban.total.area +
           Water.total.area +
           CropIrrigate.total.area +
           Grass.total.area +
           MosTreeHerb.total.area
         ) %>%
  mutate(pop_density = population/district_area,
         log_smr = log(SMR)) %>% 
  ungroup() %>% 
  group_by(District) %>%
  mutate(CropRainfed.prop = CropRainfed.total.area/district_area,
         Herbaceous.prop = Herbaceous.total.area/district_area,
         TreeShrub.prop = TreeShrub.total.area/district_area,
         MosCrop.prop = MosCrop.total.area/district_area,
         MosNatural.prop = MosNatural.total.area/district_area,
         BroadEvgClop.prop = BroadEvgClop.total.area/district_area,
         Shrub.prop = Shrub.total.area/district_area,
         ShrubEvg.prop = ShrubEvg.total.area/district_area,
         SparseVeg.prop = SparseVeg.total.area/district_area,
         FloodFresh.prop = FloodFresh.total.area/district_area,
         FloodSalt.prop = FloodSalt.total.area/district_area,
         Urban.prop = Urban.total.area/district_area,
         Water.prop = Water.total.area/district_area,
         CropIrrigate.prop = CropIrrigate.total.area/district_area,
         Grass.prop = Grass.total.area/district_area,
         MosTreeHerb.prop = MosTreeHerb.total.area/district_area) %>% 
  janitor::clean_names() %>%
  ungroup() %>%
  select(-c('x', 
            'disease', 
            'number_deaths', 
            'mortality_rates',
            'prevalence',
            'incidence_rate')) %>% 
  rename('cases' = 'number_cases') %>% 
  na.omit()
```
Observations with `NA` values were omitted from the data set

Land Cover variable discriptions:
`crop_rainfed` - Cropland, rainfed
`crop_irrigate` - Cropland, irrigated or post-flooding
`mos_crop` - Mosaic cropland (>50%) / natural vegetation (tree, shrub, herbaceous cover)(<50%) 
`mos_natural` - Mosaic natural vegetation (tree, shrub, herbaceous cover) (>50%) / cropland (<50%) 
`mos_tree_herb` - Mosaic tree and shrub (>50%) / herbaceous cover (<50%)
`herbaceous` - Herbaceous cover
`tree_shrub` - Tree or shrub cover	
`broad_evg_clop` - Tree cover, broadleaved, evergreen, closed to open (>15%)
`shrub` - Shrubland
`shrub_evg` - Shrubland evergreen	
`sparse_veg` - Sparse vegetation (tree, shrub, herbaceous cover) (<15%)	
`flood_fresh` - Tree cover, flooded, fresh or brakish water	
`flood_salt` - Tree cover, flooded, saline water	
`grass` - Grassland
`urban` - Urban areas	
`water` - Water bodies	


The Sarawak dataset does not have the following variables: 
`grass_n_patches`
`grass_patch_density`
`grass_total_area`
`grass_prop`
`mos_tree_herb_n_patches`
`mos_tree_herb_patch_density`
`mos_tree_herb_total_area`
`mos_tree_herb_prop`

Creating one dataset, if I need to run model on one
```{r}
sbh2 = sbh %>% 
  select(-c('grass_n_patches',
            'grass_patch_density',
            'grass_total_area',
            'grass_prop',
            'mos_tree_herb_n_patches',
            'mos_tree_herb_patch_density',
            'mos_tree_herb_total_area',
            'mos_tree_herb_prop'))

borneo = rbind(swk, sbh2)
```


Correlation 
```{r}
sbh_sub = sbh %>% 
  filter(year == 2000) %>% 
  select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "shrub_evg_prop", "shrub_evg_n_patches", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density","crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "shrub_evg_patch_density", "prec_mean"))

res_sub <- rcorr(as.matrix(sbh_sub))

corrplot(res_sub$r, type = "lower", 
        p.mat = res_sub$P, sig.level = 0.01, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex = 0.5)

swk_sub = swk %>%
  filter(year == 2011) %>% 
  select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "shrub_evg_prop", "shrub_evg_n_patches", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density","crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "shrub_evg_patch_density", "prec_mean"))

res_sub2 <- rcorr(as.matrix(swk_sub))

corrplot(res_sub2$r, type = "lower", 
        p.mat = res_sub2$P, sig.level = 0.01, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex=0.5)
```


```{r}
sbh.x = model.matrix(smr~., sbh_sub, na.action = NULL)[,-1]
sbh.y = sbh_sub$smr

swk.x = model.matrix(smr~., swk_sub)[,-1]
swk.y = swk_sub$smr
```


Scatterplot of predictors vs. outcome
```{r}
featurePlot(sbh.x, 
            sbh.y, 
            plot = "scatter", 
            labels = c("Predictors","SMR"),
            type = c('p'))

featurePlot(swk.x, 
            swk.y, 
            plot = "scatter", 
            labels = c("Predictors","SMR"),
            type = c('p'))
```


```{r}
sbh = sbh %>% 
  group_by(district, year) %>% 
  mutate(agri_total_area = crop_rainfed_total_area + 
           crop_irrigate_total_area +
           herbaceous_total_area +
           tree_shrub_total_area +
           mos_crop_total_area + 
           mos_natural_total_area) %>% 
  mutate(agri_prop = agri_total_area/district_area)

swk = swk %>% 
  group_by(district, year) %>% 
  mutate(agri_total_area = crop_rainfed_total_area + 
           crop_irrigate_total_area +
           herbaceous_total_area +
           tree_shrub_total_area +
           mos_crop_total_area + 
           mos_natural_total_area) %>% 
  mutate(agri_prop = agri_total_area/district_area)
```

### EDA and Visualization

Looking at the distribution of the outcome (SMR) for both Sarawak and Sabah
```{r warning=FALSE}
ggplot(data = swk, aes(x = smr)) +  
  geom_density() +
  labs(title = 'SMR distribution, Sarawak')

ggplot(data = sbh, aes(x = smr)) +  
  geom_density() +
  labs(title = 'SMR distribution, Sabah') 
```
Distribution seems to follow a poisson distribution. When running GAM, incorporate Poisson regression. Also, SMR by its nature of being a rate follows  poisson regression.

Mean precipitation across districts of Sabah and Sarawak
```{r}
ggplot(data = swk, aes(x = as.integer(year), y = prec_mean, color = district)) +  
  geom_line() +
  labs(title = 'Mean Precipitation, Sarawak')

ggplot(data = sbh, aes(x = as.integer(year), y = prec_mean, color = district)) +  
  geom_line() +
  labs(title = 'Mean Precipitation, Sabah')
```

Forest cover across districts, Sarawak and Sabah
```{r}
#Total area
ggplot(data = swk, aes(x = as.integer(year), y = broad_evg_clop_total_area, color = district)) +  
  geom_line() +
  labs(title = 'Total Forest Cover, Sarawak')

ggplot(data = sbh, aes(x = as.integer(year), y = broad_evg_clop_total_area, color = district)) +  
  geom_line() +
  labs(title = 'Total Forest Cover, Sabah')

#Number of patches
ggplot(data = swk, aes(x = as.integer(year), y = broad_evg_clop_n_patches, color = district)) +  
  geom_line() +
  labs(title = 'Number of Patches, Sarawak')

ggplot(data = sbh, aes(x = as.integer(year), y = broad_evg_clop_n_patches, color = district)) +  
  geom_line() +
  labs(title = 'Number of Patches, Sabah')

#Patch density
ggplot(data = swk, aes(x = as.integer(year), y = broad_evg_clop_patch_density, color = district)) +  
  geom_line() +
  labs(title = 'Patch Density, Sarawak')

ggplot(data = sbh, aes(x = as.integer(year), y = broad_evg_clop_patch_density, color = district)) +  
  geom_line() +
  labs(title = 'Patch Density, Sabah')

#Patch density
ggplot(data = swk, aes(x = as.integer(year), y = broad_evg_clop_prop, color = district)) +  
  geom_line() +
  labs(title = 'Proportion Forest Cover, Sarawak')

ggplot(data = sbh, aes(x = as.integer(year), y = broad_evg_clop_prop, color = district)) +  
  geom_line() +
  labs(title = 'Proportion Forest Cover, Sabah')
```

Agricultural cover across districts, Sarawak and Sabah
```{r}
#Total area
ggplot(data = swk, aes(x = as.integer(year), y = agri_total_area, color = district)) +  
  geom_line() +
  labs(title = 'Total Agricultural Cover, Sarawak')

ggplot(data = sbh, aes(x = as.integer(year), y = agri_total_area, color = district)) +  
  geom_line() +
  labs(title = 'Total Agricultural Cover, Sabah')

#Proportion
ggplot(data = swk, aes(x = as.integer(year), y = agri_prop, color = district)) +  
  geom_line() +
  labs(title = 'Proportion Agricultural Cover, Sarawak')

ggplot(data = sbh, aes(x = as.integer(year), y = agri_prop, color = district)) +  
  geom_line() +
  labs(title = 'Proportion Agricultural Cover, Sabah')
```


### Fitting Poisson Regression Model for Sabah and Sarawak

```{r}
swk.p = glm(smr ~ 
                broad_evg_clop_n_patches +
                broad_evg_clop_patch_density +
                broad_evg_clop_prop +
                crop_irrigate_n_patches +
                crop_irrigate_patch_density +
                crop_rainfed_n_patches + 
                crop_rainfed_patch_density +
                flood_fresh_n_patches +
                flood_fresh_patch_density +
                flood_fresh_prop +
                flood_salt_n_patches +
                flood_salt_patch_density +
                flood_salt_prop +
                herbaceous_n_patches +
                herbaceous_patch_density +
                mos_crop_n_patches +
                mos_crop_patch_density +
                mos_natural_n_patches +
                mos_natural_patch_density +
                shrub_n_patches +
                shrub_patch_density +
                shrub_prop +
                shrub_evg_n_patches +
                shrub_evg_patch_density +
                shrub_evg_prop +
                sparse_veg_n_patches +
                sparse_veg_patch_density +
                sparse_veg_prop +
                tree_shrub_n_patches +
                tree_shrub_patch_density + 
                urban_n_patches +
                urban_patch_density +
                urban_prop +
                water_n_patches +
                water_patch_density +
                agri_prop +
                water_prop +
                pop_density +
                prec_mean,
              data = swk, family = poisson(link = log)
              )

summary(swk.p)
plot(swk.p)
```


```{r}
swk.lm = lm(smr ~ 
                broad_evg_clop_n_patches +
                broad_evg_clop_patch_density +
                broad_evg_clop_prop +
                crop_irrigate_n_patches +
                crop_irrigate_patch_density +
                crop_rainfed_n_patches + 
                crop_rainfed_patch_density +
                flood_fresh_n_patches +
                flood_fresh_patch_density +
                flood_fresh_prop +
                flood_salt_n_patches +
                flood_salt_patch_density +
                flood_salt_prop +
                herbaceous_n_patches +
                herbaceous_patch_density +
                mos_crop_n_patches +
                mos_crop_patch_density +
                mos_natural_n_patches +
                mos_natural_patch_density +
                shrub_n_patches +
                shrub_patch_density +
                shrub_prop +
                shrub_evg_n_patches +
                #shrub_evg_patch_density +
                #shrub_evg_prop +
                sparse_veg_n_patches +
                sparse_veg_patch_density +
                sparse_veg_prop +
                tree_shrub_n_patches +
                tree_shrub_patch_density + 
                urban_n_patches +
                urban_patch_density +
                urban_prop +
                water_n_patches +
                water_patch_density +
                agri_prop +
                #water_prop +
                pop_density +
                prec_mean,
              data = swk
              )

summary(swk.lm)
plot(swk.lm)
```


Checking variance and mean for both datasets
```{r}
labels = c('Region', 'Mean', 'Variance')
Sabah = c(mean(sbh$smr, na.rm = T), var(sbh$smr, na.rm = T))
Sarawak = c(mean(swk$smr, na.rm = T), var(swk$smr, na.rm = T))

mean_and_variance = rbind(Sabah, Sarawak)
```
For both regions the mean and variance differ, indicating the need to use a negative binomal distribution instead of poisson.

### Spatial Stuff
```{r}
library(raster)
library(spdep)
malaysia = getData('GADM', country='MYS', level=2)
sabah = malaysia[malaysia$NAME_1 == 'Sabah',]
sarawak = malaysia[malaysia$NAME_1 == 'Sarawak',]

df = droplevels(as(sabah, 'data.frame'))

aea.proj = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96"
shp = spTransform(sabah, CRS(aea.proj))  # project to Albers
shpf = fortify(shp, region = 'NAME_2')

df = transform(df, smr = sbh$smr)

nb <- poly2nb(shp, row.names = df$NAME_2)
names(nb) <- attr(nb, "region.id")
str(nb[1:6])

m1 <- gam(smr ~ s(NAME_2, bs = 'mrf', xt = list(nb = nb)), # define MRF smooth
          data = df,
          method = 'REML')
```


### Fitting GAM for Sabah and Sarawak

splitting data into meaniful groups in order to fit data with GAMs
```{r}
train_sbh = sbh %>% 
  filter(year %in% c(2000, 2001, 2002, 2003, 2005, 2006, 2008, 2009, 2010))

test_sbh = sbh %>% 
  filter(year %in% c(2011, 2012, 2013)) 
```

```{r}
sbh_forest.fit = gamm(smr ~ te(broad_evg_clop_prop, year, k = c(20, 9)) +
                       te(broad_evg_clop_n_patches, year, k = c(10, 9)) +
                       te(broad_evg_clop_patch_density, year, k = c(10, 9)),
                     data = train_sbh, method = 'REML')

sbh_agri.fit = gam(smr ~ te(agri_prop, year) +
                       te(crop_rainfed_n_patches, year) +
                       te(crop_rainfed_patch_density, year) +
                       te(crop_irrigate_n_patches, year) +
                       te(crop_irrigate_patch_density, year) +
                       te(mos_crop_n_patches, year) +
                       te(mos_crop_patch_density, year) +
                       te(mos_natural_n_patches, year) +
                       te(mos_natural_patch_density, year),
                     data = train_sbh, method = 'REML')

sbh_water.fit = gam(smr ~ te(prec_mean, year) +
                      te(tree_shrub_patch_density, year) +
                      te(flood_fresh_n_patches, year) +
                      te(flood_fresh_patch_density, year) +
                      te(flood_fresh_prop, year) +
                      te(flood_salt_n_patches, year) +
                      te(flood_salt_patch_density, year) +
                      te(flood_salt_prop, year),
                    data = train_sbh, method = 'REML')

#plot(sbh_forest.fit$gam)
#summary(sbh_forest.fit$gam)
#gam.check(sbh_forest.fit$gam)

pred_forest = predict(sbh_forest.fit$gam, newdata = test_sbh)
pred_agri = predict(sbh_agri.fit, newdata = test_sbh)
pred_water = predict(sbh_water.fit, newdata = test_sbh)

mean(data.matrix(test_sbh[,57] - pred_forest)^2)
mean(data.matrix(test_sbh[,57] - pred_agri)^2)
mean(data.matrix(test_sbh[,57] - pred_water)^2)

```


Sarawak
```{r}
m = gamm(smr ~ s(year, k = 12),
              data = sbh, method = 'REML')

summary(m)
plot(m$gam)
```

```{r}
layout(matrix(1:2, ncol = 2))
acf(resid(m$lme), lag.max = 36, main = "ACF")
pacf(resid(m$lme), lag.max = 36, main = "pACF")
layout(1)
```


Autocorrelation for time (year)
```{r}
ctrl <- list(niterEM = 0, msVerbose = TRUE, optimMethod="L-BFGS-B")

#AR(1)
m1 = gamm(smr ~ s(year, k = 12),
           data = sbh, method = 'REML', correlation = corARMA(form = ~ 1|year, p = 1),
           control = ctrl)
## AR(2)
m2 = gamm(smr ~ s(year, k = 12),
           data = sbh, method = 'REML', correlation = corARMA(form = ~ 1|year, p = 2),
           control = ctrl)

## AR(3)
m3 = gamm(smr ~ s(year, k = 12),
           data = sbh, method = 'REML', correlation = corARMA(form = ~ 1|year, p = 3),
           control = ctrl)
```

```{r}
anova(m$lme, m1$lme, m2$lme, m3$lme)
```



```{r}
 = gam(smr ~ s(prec_mean) +
                s(broad_evg_clop_prop) +
                s(broad_evg_clop_n_patches) +
                s(broad_evg_clop_patch_density) +
                s(crop_irrigate_n_patches) +
                s(crop_irrigate_patch_density) +
                s(crop_rainfed_n_patches) +
                s(crop_rainfed_patch_density) +
                s(herbaceous_n_patches) +
                s(herbaceous_patch_density) +
                s(mos_crop_n_patches) +
                s(mos_crop_patch_density) +
                s(mos_natural_n_patches) +
                s(mos_natural_patch_density) +
                s(tree_shrub_n_patches) +
                s(tree_shrub_patch_density) +
                s(flood_fresh_n_patches) +
                s(flood_fresh_patch_density) +
                s(flood_fresh_prop) +
                s(flood_salt_n_patches) +
                s(flood_salt_patch_density) +
                s(flood_salt_prop) +
                s(shrub_n_patches) +
                s(shrub_patch_density) +
                s(shrub_prop) +
                s(shrub_evg_n_patches) +
                s(shrub_evg_patch_density) +
                s(shrub_evg_prop) +
                s(sparse_veg_n_patches) +
                s(sparse_veg_patch_density) +
                s(sparse_veg_prop) +
                s(grass_n_patches) +
                s(grass_patch_density) +
                s(grass_prop) +
                s(urban_n_patches) +
                s(urban_patch_density) +
                s(urban_prop) +
                s(water_n_patches) +
                s(water_patch_density) +
                s(water_prop) +
                s(year) +
                district
                  ,
              data = sbh)

summary(sbh.fit)
gam.check(sbh.fit)
plot(sbh.fit)
```

### Using Caret

preprocess data
```{r}
sbh_var = sbh %>% 
  select(c("smr", 
           "broad_evg_clop_prop", 
           "crop_rainfed_prop", 
           "herbaceous_prop", 
           "shrub_prop", "urban_prop", 
           "tree_shrub_prop", 
           "mos_crop_prop", 
           "mos_natural_prop", 
           "sparse_veg_prop", 
           "flood_fresh_prop", 
           "flood_salt_prop", 
           "water_prop", 
           "crop_irrigate_prop", 
           "shrub_evg_prop", 
           "shrub_evg_n_patches", 
           "broad_evg_clop_n_patches", 
           "crop_rainfed_n_patches", 
           "herbaceous_n_patches", 
           "shrub_n_patches", 
           "urban_n_patches", 
           "tree_shrub_n_patches", 
           "mos_crop_n_patches", 
           "mos_natural_n_patches", 
           "sparse_veg_n_patches", 
           "flood_fresh_n_patches", 
           "flood_salt_n_patches", 
           "water_n_patches", 
           "crop_irrigate_n_patches", 
           "broad_evg_clop_patch_density",
           "crop_rainfed_patch_density", 
           "herbaceous_patch_density", 
           "shrub_patch_density", 
           "urban_patch_density", 
           "tree_shrub_patch_density", 
           "mos_crop_patch_density", 
           "mos_natural_patch_density", 
           "sparse_veg_patch_density", 
           "flood_fresh_patch_density", 
           "flood_salt_patch_density", 
           "water_patch_density", 
           "crop_irrigate_patch_density", 
           "shrub_evg_patch_density",
           "grass_n_patches",
           "grass_patch_density",
           "grass_prop",
           "mos_tree_herb_n_patches",
           "mos_tree_herb_patch_density",
           "mos_tree_herb_prop",
           "prec_mean", 
           "year", 
           "district"))


swk_var = swk %>%
  select(c("smr", 
           "broad_evg_clop_prop", 
           "crop_rainfed_prop", 
           "herbaceous_prop", 
           "shrub_prop", "urban_prop", 
           "tree_shrub_prop", 
           "mos_crop_prop", 
           "mos_natural_prop", 
           "sparse_veg_prop", 
           "flood_fresh_prop", 
           "flood_salt_prop", 
           "water_prop", 
           "crop_irrigate_prop", 
           "shrub_evg_prop", 
           "shrub_evg_n_patches", 
           "broad_evg_clop_n_patches", 
           "crop_rainfed_n_patches", 
           "herbaceous_n_patches", 
           "shrub_n_patches", 
           "urban_n_patches", 
           "tree_shrub_n_patches", 
           "mos_crop_n_patches", 
           "mos_natural_n_patches", 
           "sparse_veg_n_patches", 
           "flood_fresh_n_patches", 
           "flood_salt_n_patches", 
           "water_n_patches", 
           "crop_irrigate_n_patches", 
           "broad_evg_clop_patch_density",
           "crop_rainfed_patch_density", 
           "herbaceous_patch_density", 
           "shrub_patch_density", 
           "urban_patch_density", 
           "tree_shrub_patch_density", 
           "mos_crop_patch_density", 
           "mos_natural_patch_density", 
           "sparse_veg_patch_density", 
           "flood_fresh_patch_density", 
           "flood_salt_patch_density", 
           "water_patch_density", 
           "crop_irrigate_patch_density", 
           "shrub_evg_patch_density", 
           "prec_mean", 
           "year", 
           "district"))

sbh.x = model.matrix(smr~., sbh_var, na.action = NULL)[,-1]
sbh.y = sbh_sub$smr

swk.x = model.matrix(smr~., swk_var)[,-1]
swk.y = swk_sub$smr

# add interaction terms
sbh.yd = as.data.frame(sbh.x) %>% 
  select("year2001":"districtTuaran") 
sbh.yd = as.matrix(sbh.yd)

sbh.temp = sbh_var %>% 
  select(-c("year", "district")) 
sbh.temp = as.matrix(sbh.temp)

sbh.int = sbh.yd*sbh.temp
```


sbh
```{r}
ctrl1 <- trainControl(method = "cv", number = 10)

set.seed(1)

sbh.lm = train(sbh.x, sbh.y,
                 method = "lm",
                 preProc = c("scale"), #scaling data in order to compare coefficients
                 trControl = ctrl1)

var_imp = as.data.frame(sbh.lm$finalModel$coefficients)

set.seed(1)

swk.lm = train(swk.x, swk.y,
                 method = "lm",
                 preProc = c("scale"), #scaling data in order to compare coefficients
                 trControl = ctrl1)

var_imp_swk = as.data.frame(swk.lm$finalModel$coefficients)
```



Inspect model:
gam.check()
concurvity()
VIF/tolerance
summary()
coef()

