---
title: "Analysis"
author: "Matthew Perrotta"
date: "April 20, 2019"
output: html_document
---

Data analysis for MPH Epidemiology thesis requirement.

Research Question: What is the best predictor of standardized malaria rate (SMR) in the Malaysian regions of Sarawak and Sabah.

Methods:
-Exploratory data analysis and visualizations
-multiple linear regression
-more advanced linear regression methods
-GAM

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Libraries
```{r echo = T, message = FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(mgcv)
library(corrplot)
library(Hmisc)
library(caret)
library(PerformanceAnalytics)
```

READ: https://www.nature.com/articles/s41598-017-02560-z

### Import and clean data
```{r}
swk = read.csv('./data/data_full_sar.csv') %>% 
  group_by(district, year) %>% 
  mutate(district_area = 
           CropRainfed.total.area + 
           Herbaceous.total.area + 
           TreeShrub.total.area +
           MosCrop.total.area +
           MosNatural.total.area + 
           BroadEvgClop.total.area +
           Shrub.total.area +
           ShrubEvg.total.area +
           SparseVeg.total.area +
           FloodFresh.total.area +
           FloodSalt.total.area +
           Urban.total.area +
           Water.total.area +
           CropIrrigate.total.area 
         ) %>% 
  ungroup() %>% 
  group_by(district) %>%
  mutate(CropRainfed.prop = CropRainfed.total.area/district_area,
         Herbaceous.prop = Herbaceous.total.area/district_area,
         TreeShrub.prop = TreeShrub.total.area/district_area,
         MosCrop.prop = MosCrop.total.area/district_area,
         MosNatural.prop = MosNatural.total.area/district_area,
         BroadEvgClop.prop = BroadEvgClop.total.area/district_area,
         Shrub.prop = Shrub.total.area/district_area,
         ShrubEvg.prop = ShrubEvg.total.area/district_area,
         SparseVeg.prop = SparseVeg.total.area/district_area,
         FloodFresh.prop = FloodFresh.total.area/district_area,
         FloodSalt.prop = FloodSalt.total.area/district_area,
         Urban.prop = Urban.total.area/district_area,
         Water.prop = Water.total.area/district_area,
         CropIrrigate.prop = CropIrrigate.total.area/district_area) %>% 
  janitor::clean_names() %>%
  ungroup() %>%  
  select(-c('x',
            'population_year',
            'cases_year',
            'expected',
            'sd')) %>% 
  rename('smr' = 'sir',
         'prec_mean' = 'mean',
         'cases' = 'case_number') %>% 
  mutate(year = as.factor(year),
         log_smr = log(smr))


sbh = read.csv('./data/dataFull.csv') %>% 
  group_by(District, Year) %>% 
  mutate(district_area = 
           CropRainfed.total.area + 
           Herbaceous.total.area + 
           TreeShrub.total.area +
           MosCrop.total.area +
           MosNatural.total.area + 
           BroadEvgClop.total.area +
           Shrub.total.area +
           ShrubEvg.total.area +
           SparseVeg.total.area +
           FloodFresh.total.area +
           FloodSalt.total.area +
           Urban.total.area +
           Water.total.area +
           CropIrrigate.total.area +
           Grass.total.area +
           MosTreeHerb.total.area
         ) %>% 
  ungroup() %>% 
  group_by(District) %>%
  mutate(CropRainfed.prop = CropRainfed.total.area/district_area,
         Herbaceous.prop = Herbaceous.total.area/district_area,
         TreeShrub.prop = TreeShrub.total.area/district_area,
         MosCrop.prop = MosCrop.total.area/district_area,
         MosNatural.prop = MosNatural.total.area/district_area,
         BroadEvgClop.prop = BroadEvgClop.total.area/district_area,
         Shrub.prop = Shrub.total.area/district_area,
         ShrubEvg.prop = ShrubEvg.total.area/district_area,
         SparseVeg.prop = SparseVeg.total.area/district_area,
         FloodFresh.prop = FloodFresh.total.area/district_area,
         FloodSalt.prop = FloodSalt.total.area/district_area,
         Urban.prop = Urban.total.area/district_area,
         Water.prop = Water.total.area/district_area,
         CropIrrigate.prop = CropIrrigate.total.area/district_area,
         Grass.prop = Grass.total.area/district_area,
         MosTreeHerb.prop = MosTreeHerb.total.area/district_area) %>% 
  janitor::clean_names() %>%
  ungroup() %>%
  select(-c('x', 
            'disease', 
            'number_deaths', 
            'mortality_rates',
            'prevalence',
            'incidence_rate')) %>% 
  rename('cases' = 'number_cases') %>% 
  mutate(year = as.factor(year),
         log_smr = log(smr))
```

Land Cover variable discriptions:
`crop_rainfed` - Cropland, rainfed
`crop_irrigate` - Cropland, irrigated or post-flooding
`mos_crop` - Mosaic cropland (>50%) / natural vegetation (tree, shrub, herbaceous cover)(<50%) 
`mos_natural` - Mosaic natural vegetation (tree, shrub, herbaceous cover) (>50%) / cropland (<50%) 
`mos_tree_herb` - Mosaic tree and shrub (>50%) / herbaceous cover (<50%)
`herbaceous` - Herbaceous cover
`tree_shrub` - Tree or shrub cover	
`broad_evg_clop` - Tree cover, broadleaved, evergreen, closed to open (>15%)
`shrub` - Shrubland
`shrub_evg` - Shrubland evergreen	
`sparse_veg` - Sparse vegetation (tree, shrub, herbaceous cover) (<15%)	
`flood_fresh` - Tree cover, flooded, fresh or brakish water	
`flood_salt` - Tree cover, flooded, saline water	
`grass` - Grassland
`urban` - Urban areas	
`water` - Water bodies	


The Sarawak dataset does not have the following variables: 
grass_n_patches
grass_patch_density
grass_total_area
grass_prop
mos_tree_herb_n_patches
mos_tree_herb_patch_density
mos_tree_herb_total_area
mos_tree_herb_prop

I will remove these variables from Sabah
```{r}
sbh = sbh %>% 
  select(-c('grass_n_patches', 
            'grass_patch_density', 
            'grass_total_area', 
            'grass_prop', 
            'mos_tree_herb_n_patches', 
            'mos_tree_herb_patch_density', 
            'mos_tree_herb_total_area', 
            'mos_tree_herb_prop'))
```



### EDA and Visualization

Looking at the distribution of the outcome (SMR) for both Sarawak and Sabah
```{r warning=FALSE}
ggplot(data = swk, aes(x = smr)) +  
  geom_histogram() +
  labs(title = 'SMR distribution, Sarawak')

ggplot(data = sbh, aes(x = smr)) +  
  geom_histogram() +
  labs(title = 'SMR distribution, Sabah') 
```
Distribution seems to follow a poisson distribution. When running GAM, incorporate Poisson regression. Also, SMR by its nature of being a rate follows  poisson regression.


Correlation 
```{r}
sbh_sub = sbh %>% 
  select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density","crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "prec_mean"))

res_sub <- rcorr(as.matrix(sbh_sub))

corrplot(res_sub$r, type = "lower", 
        p.mat = res_sub$P, sig.level = 0.01, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex = 0.5)

swk_sub = subset(swk, select = c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density","crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "prec_mean"))

res_sub2 <- rcorr(as.matrix(swk_sub))

corrplot(res_sub2$r, type = "lower", 
        p.mat = res_sub2$P, sig.level = 0.01, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex=0.5)
```


```{r}
sbh.x = model.matrix(smr~., sbh_sub, na.action = NULL)[,-1]
sbh.y = sbh_sub$smr

swk.x = model.matrix(smr~., swk_sub)[,-1]
swk.y = swk_sub$smr
```


Scatterplot of predictors vs. outcome
```{r}
featurePlot(sbh.x, 
            sbh.y, 
            plot = "scatter", 
            labels = c("Predictors","SMR"),
            type = c('p'))
```


```{r}
swk %>% 
  mutate(district = fct_reorder(district, smr)) %>% 
  ggplot() +  
  geom_point(aes(x = district,y = smr)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

swk %>% 
  #mutate(district = fct_reorder(year, smr)) %>% 
  ggplot() +  
  geom_boxplot(aes(x = year,y = smr)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(data = swk) +
  geom_point(aes(x = prec_mean, y = log(smr)))

ggplot(data = swk) +
  geom_point(aes(x = broad_evg_clop_total_area, y = log(smr)))
  
  
sbh.glm = glm(smr ~ urban_n_patches, data = sbh, family = poisson(link = log))
summary(sbh.glm)
plot(sbh.glm)
```


### Fitting GAM for SABAH

GAM, without Poisson
```{r}
sbh.fit1 = bam(smr ~ 
                s(broad_evg_clop_n_patches, k = 5) +
                s(broad_evg_clop_patch_density, k = 5) +
                s(broad_evg_clop_prop, k = 5) +
                s(crop_irrigate_n_patches, k = 5) +
                s(crop_irrigate_patch_density, k = 5) +
                s(crop_irrigate_prop, k = 5) +
                s(crop_rainfed_n_patches, k = 5) + 
                s(crop_rainfed_patch_density, k = 5) +
                s(crop_rainfed_prop, k = 5) +
                s(flood_fresh_n_patches, k = 5) +
                s(flood_fresh_patch_density, k = 5) +
                s(flood_fresh_prop, k = 5) +
                s(flood_salt_n_patches, k = 5) +
                s(flood_salt_patch_density, k = 5) +
                s(flood_salt_prop, k = 5) +
                s(herbaceous_n_patches, k = 5) +
                s(herbaceous_patch_density, k = 5) +
                s(herbaceous_prop, k = 5) +
                s(mos_crop_n_patches, k = 5) +
                s(mos_crop_patch_density, k = 5) +
                s(mos_crop_prop, k = 5) +
                s(mos_natural_n_patches, k = 5) +
                s(mos_natural_patch_density, k = 5) +
                s(mos_natural_prop, k = 5) +
                s(shrub_n_patches, k = 5) +
                s(shrub_patch_density, k = 5) +
                s(shrub_prop, k = 5) +
                s(sparse_veg_n_patches, k = 5) +
                s(sparse_veg_patch_density, k = 5) +
                s(sparse_veg_prop, k = 5) +
                s(tree_shrub_n_patches, k = 5) +
                s(tree_shrub_patch_density, k = 5) + 
                s(tree_shrub_prop, k = 5) +
                s(urban_n_patches, k = 5) +
                s(urban_patch_density, k = 5) +
                s(urban_prop, k = 5) +
                s(water_n_patches, k = 5) +
                s(water_patch_density, k = 5) +
                s(water_prop, k = 5) +
                s(prec_mean, k = 5) +
                year +
                district, 
              data = sbh, method = 'REML',
              #family = poisson(link = log)
              ) 

summary(sbh.fit1)
gam.check(sbh.fit1)
```

GAM, with Poisson
```{r}
sbh.fit1 = bam(smr ~ 
                #s(broad_evg_clop_n_patches, k = 5) +
                s(broad_evg_clop_patch_density, k = 5) +
                s(broad_evg_clop_prop, k = 5) +
                #s(crop_irrigate_n_patches, k = 5) +
                s(crop_irrigate_patch_density, k = 5) +
                s(crop_irrigate_prop, k = 5) +
                #s(crop_rainfed_n_patches, k = 5) + 
                s(crop_rainfed_patch_density, k = 5) +
                s(crop_rainfed_prop, k = 5) +
                #s(flood_fresh_n_patches, k = 5) +
                s(flood_fresh_patch_density, k = 5) +
                s(flood_fresh_prop, k = 5) +
                #s(flood_salt_n_patches, k = 5) +
                s(flood_salt_patch_density, k = 5) +
                s(flood_salt_prop, k = 5) +
                #s(herbaceous_n_patches, k = 5) +
                s(herbaceous_patch_density, k = 5) +
                s(herbaceous_prop, k = 5) +
                #s(mos_crop_n_patches, k = 5) +
                s(mos_crop_patch_density, k = 5) +
                s(mos_crop_prop, k = 5) +
                #s(mos_natural_n_patches, k = 5) +
                s(mos_natural_patch_density, k = 5) +
                s(mos_natural_prop, k = 5) +
                #s(shrub_n_patches, k = 5) +
                s(shrub_patch_density, k = 5) +
                s(shrub_prop, k = 5) +
                #s(sparse_veg_n_patches, k = 5) +
                s(sparse_veg_patch_density, k = 5) +
                s(sparse_veg_prop, k = 5) +
                #s(tree_shrub_n_patches, k = 5) +
                s(tree_shrub_patch_density, k = 5) + 
                s(tree_shrub_prop, k = 5) +
                #s(urban_n_patches, k = 5) +
                s(urban_patch_density, k = 5) +
                s(urban_prop, k = 5) +
                #s(water_n_patches, k = 5) +
                s(water_patch_density, k = 5) +
                s(water_prop, k = 5) +
                s(prec_mean, k = 5) +
                year +
                district, 
              data = sbh, method = 'REML',
              family = poisson(link = log)
              ) 

summary(sbh.fit2)
```

Inspect model:
gam.check()
concurvity()
VIF/tolerance
summary()
coef()
```{r}
summary(sbh_fit)
gam.check(gam_sbh)
```

