---
title: "Exploration of Data"
author: "Matthew Perrotta"
date: "5/10/2019"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Libraries
```{r echo = T, message = FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(mgcv)
library(corrplot)
library(Hmisc)
library(caret)
library(PerformanceAnalytics)
library(car)
library(olsrr)
#library(MASS)
#library(raster)
```

READ: https://www.nature.com/articles/s41598-017-02560-z

# Import and clean data

Two datasets are created, `swk` for Sarawak and `sbh` for Sabah. Variables that are not needed are removed.
```{r}
swk = read.csv('./data/data_full_sar.csv') %>% 
  group_by(district, year) %>% 
  mutate(district_area = 
           CropRainfed.total.area + 
           Herbaceous.total.area + 
           TreeShrub.total.area +
           MosCrop.total.area +
           MosNatural.total.area + 
           BroadEvgClop.total.area +
           Shrub.total.area +
           ShrubEvg.total.area +
           SparseVeg.total.area +
           FloodFresh.total.area +
           FloodSalt.total.area +
           Urban.total.area +
           Water.total.area +
           CropIrrigate.total.area 
         ) %>%
  ungroup() %>% 
  group_by(district) %>%
  mutate(CropRainfed.prop = CropRainfed.total.area/district_area,
         Herbaceous.prop = Herbaceous.total.area/district_area,
         TreeShrub.prop = TreeShrub.total.area/district_area,
         MosCrop.prop = MosCrop.total.area/district_area,
         MosNatural.prop = MosNatural.total.area/district_area,
         BroadEvgClop.prop = BroadEvgClop.total.area/district_area,
         Shrub.prop = Shrub.total.area/district_area,
         ShrubEvg.prop = ShrubEvg.total.area/district_area,
         SparseVeg.prop = SparseVeg.total.area/district_area,
         FloodFresh.prop = FloodFresh.total.area/district_area,
         FloodSalt.prop = FloodSalt.total.area/district_area,
         Urban.prop = Urban.total.area/district_area,
         Water.prop = Water.total.area/district_area,
         CropIrrigate.prop = CropIrrigate.total.area/district_area) %>% 
  janitor::clean_names() %>%
  ungroup() %>%  
  dplyr::select(-c('x',
            'population_year',
            'cases_year',
            'expected',
            'sd')) %>% 
  rename('smr' = 'sir',
         'prec_mean' = 'mean',
         'cases' = 'case_number') %>% 
  mutate(district = str_replace(district, '_', ' ')) %>% 
  na.omit()


sbh = read.csv('./data/dataFull.csv') %>% 
  group_by(District, Year) %>% 
  mutate(district_area = 
           CropRainfed.total.area + 
           Herbaceous.total.area + 
           TreeShrub.total.area +
           MosCrop.total.area +
           MosNatural.total.area + 
           BroadEvgClop.total.area +
           Shrub.total.area +
           ShrubEvg.total.area +
           SparseVeg.total.area +
           FloodFresh.total.area +
           FloodSalt.total.area +
           Urban.total.area +
           Water.total.area +
           CropIrrigate.total.area +
           Grass.total.area +
           MosTreeHerb.total.area
         ) %>%
  ungroup() %>% 
  group_by(District) %>%
  mutate(CropRainfed.prop = CropRainfed.total.area/district_area,
         Herbaceous.prop = Herbaceous.total.area/district_area,
         TreeShrub.prop = TreeShrub.total.area/district_area,
         MosCrop.prop = MosCrop.total.area/district_area,
         MosNatural.prop = MosNatural.total.area/district_area,
         BroadEvgClop.prop = BroadEvgClop.total.area/district_area,
         Shrub.prop = Shrub.total.area/district_area,
         ShrubEvg.prop = ShrubEvg.total.area/district_area,
         SparseVeg.prop = SparseVeg.total.area/district_area,
         FloodFresh.prop = FloodFresh.total.area/district_area,
         FloodSalt.prop = FloodSalt.total.area/district_area,
         Urban.prop = Urban.total.area/district_area,
         Water.prop = Water.total.area/district_area,
         CropIrrigate.prop = CropIrrigate.total.area/district_area,
         Grass.prop = Grass.total.area/district_area,
         MosTreeHerb.prop = MosTreeHerb.total.area/district_area) %>% 
  janitor::clean_names() %>%
  ungroup() %>%
  select(-c('x', 
            'disease', 
            'number_deaths', 
            'mortality_rates',
            'prevalence',
            'incidence_rate')) %>% 
  rename('cases' = 'number_cases') %>% 
  na.omit()
```
Observations with `NA` values were omitted from the data set

Land Cover variable discriptions:

  * `crop_rainfed` - Cropland, rainfed
  * `crop_irrigate` - Cropland, irrigated or post-flooding
  * `mos_crop` - Mosaic cropland (>50%) / natural vegetation (tree, shrub, herbaceous cover)(<50%) 
  * `mos_natural` - Mosaic natural vegetation (tree, shrub, herbaceous cover) (>50%) / cropland (<50%) 
  * `mos_tree_herb` - Mosaic tree and shrub (>50%) / herbaceous cover (<50%)
  * `herbaceous` - Herbaceous cover
  * `tree_shrub` - Tree or shrub cover	
  * `broad_evg_clop` - Tree cover, broadleaved, evergreen, closed to open (>15%)
  * `shrub` - Shrubland
  * `shrub_evg` - Shrubland evergreen	
  * `sparse_veg` - Sparse vegetation (tree, shrub, herbaceous cover) (<15%)	
  * `flood_fresh` - Tree cover, flooded, fresh or brakish water	
  * `flood_salt` - Tree cover, flooded, saline water	
  * `grass` - Grassland
  * `urban` - Urban areas	
  * `water` - Water bodies	


The Sarawak dataset does not have the following variables: 

  * `grass_n_patches`
  * `grass_patch_density`
  * `grass_total_area`
  * `grass_prop`
  * `mos_tree_herb_n_patches`
  * `mos_tree_herb_patch_density`
  * `mos_tree_herb_total_area`
  * `mos_tree_herb_prop`

Creating a variable for agriculture
```{r}
sbh = sbh %>% 
  group_by(district, year) %>% 
  mutate(agri_total_area = crop_rainfed_total_area + 
           crop_irrigate_total_area +
           herbaceous_total_area +
           tree_shrub_total_area +
           mos_crop_total_area + 
           mos_natural_total_area) %>% 
  mutate(agri_prop = agri_total_area/district_area) %>% 
  ungroup()

swk = swk %>% 
  group_by(district, year) %>% 
  mutate(agri_total_area = crop_rainfed_total_area + 
           crop_irrigate_total_area +
           herbaceous_total_area +
           tree_shrub_total_area +
           mos_crop_total_area + 
           mos_natural_total_area) %>% 
  mutate(agri_prop = agri_total_area/district_area) %>% 
  ungroup()
```


# Exploration of the Data


```{r}
sbh_sub = sbh %>% 
  select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "shrub_evg_prop", "shrub_evg_n_patches", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density","crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "shrub_evg_patch_density", "prec_mean"))

swk_sub = swk %>% 
  select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "shrub_evg_prop", "shrub_evg_n_patches", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density","crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "shrub_evg_patch_density", "prec_mean"))

sbh.x = model.matrix(smr~., sbh_sub, na.action = NULL)[,-1]
sbh.y = sbh_sub$smr

swk.x = model.matrix(smr~., swk_sub)[,-1]
swk.y = swk_sub$smr

featurePlot(sbh.x, 
            sbh.y, 
            plot = "scatter", 
            labels = c("Predictors","SMR"),
            type = c('p'))

featurePlot(swk.x, 
            swk.y, 
            plot = "scatter", 
            labels = c("Predictors","SMR"),
            type = c('p'))
```

```{r}
library(graphics)
library(gridExtra)
# Plots 
#par(mfrow=c(1,2))
# boxplot(SMR ~ Year, xlab = "Year",
#         ylab = "Standarized Mortality Ratio (SMR)",connect=T, data=data_full, outline=F, col = "lightgray")
# boxplot(SMR ~ District, xlab = "Districts",
#         ylab = "Standarized Mortality Ratio (SMR)",connect=T, data=data_full, outline=F, col = "lightgray")
plot1 <- ggplot(data = sbh, aes(x = district, y = smr, group = district)) +  
  geom_boxplot(outlier.shape=NA ) +  
  ylim(0, 6.5) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y="SMR, Sabah")
   
plot2 <- ggplot(data = sbh, aes(x = year, y = smr, group = year)) +  
  geom_boxplot(outlier.shape=NA  ) + ylim(0, 6.5) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(y="SMR, Sabah")
require(cowplot)
plot_grid(plot1, plot2, align='h', labels=c('', ''))
```

Sarawak
```{r}
library(graphics)
library(gridExtra)

plot1 <- ggplot(data = swk, aes(x = district, y = smr, group = district)) +  
  geom_boxplot(outlier.shape=NA ) +  
  ylim(0, 6.5) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y="SMR, Sarawak")
   
plot2 <- ggplot(data = swk, aes(x = year, y = smr, group = year)) +  
  geom_boxplot(outlier.shape=NA  ) + ylim(0, 6.5) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(y="SMR, Sarawak")
require(cowplot)
plot_grid(plot1, plot2, align='h', labels=c('', ''))
```


Distribution of SMR
```{r warning=FALSE}
ggplot(data = swk, aes(x = smr)) +  
  geom_histogram() +
  labs(title = 'SMR distribution, Sarawak')

ggplot(data = sbh, aes(x = smr)) +  
  geom_histogram() +
  labs(title = 'SMR distribution, Sabah') 
```

Distribution of cases
```{r warning=FALSE}
ggplot(data = swk, aes(x = cases)) +  
  geom_histogram() +
  labs(title = 'Distribution of Cases, Sarawak')

ggplot(data = sbh, aes(x = cases)) +  
  geom_histogram() +
  labs(title = 'Distribution of Cases, Sabah') 
```

Cases overtime
```{r}
swk %>% 
  group_by(year) %>% 
  mutate(total_cases = sum(cases)) %>% 
  ggplot(aes(y = total_cases, x = year)) +
  geom_line() +
  labs(title = 'Cases overtime, Sarawak')

sbh %>% 
  group_by(year) %>% 
  mutate(total_cases = sum(cases)) %>% 
  ggplot(aes(y = total_cases, x = year)) +
  geom_line() +
  labs(title = 'Cases overtime, Sabah')
```

# Spaghetti Plots

SMR by District
```{r}
swk %>%
  ggplot(aes(y = smr, x = year, color = district)) +
  geom_line() +
  labs(title = 'SMR by district, Sarawak')

sbh %>% 
  ggplot(aes(y = smr, x = year, color = district)) +
  geom_line() +
  labs(title = 'SMR by district, Sabah')
```

Cases by district
```{r}
swk %>%
  ggplot(aes(y = cases, x = year, color = district)) +
  geom_line() +
  labs(title = 'Cases by district, Sarawak')

sbh %>% 
  ggplot(aes(y = cases, x = year, color = district)) +
  geom_line() +
  labs(title = 'Cases by district, Sabah')
```

Mean precipitation across districts of Sabah and Sarawak
```{r}
ggplot(data = swk, aes(x = year, y = prec_mean, color = district)) +  
  geom_line() +
  labs(title = 'Mean Precipitation, Sarawak')

ggplot(data = sbh, aes(x = year, y = prec_mean, color = district)) +  
  geom_line() +
  labs(title = 'Mean Precipitation, Sabah')
```

Forest cover across districts, Sarawak and Sabah
```{r}
#Total area
ggplot(data = swk, aes(x = year, y = broad_evg_clop_total_area, color = district)) +  
  geom_line() +
  labs(title = 'Total Forest Cover, Sarawak')

ggplot(data = sbh, aes(x = year, y = broad_evg_clop_total_area, color = district)) +  
  geom_line() +
  labs(title = 'Total Forest Cover, Sabah')

#Number of patches
ggplot(data = swk, aes(x = year, y = broad_evg_clop_n_patches, color = district)) +  
  geom_line() +
  labs(title = 'Number of Patches, Sarawak')

ggplot(data = sbh, aes(x = year, y = broad_evg_clop_n_patches, color = district)) +  
  geom_line() +
  labs(title = 'Number of Patches, Sabah')

#Patch density
ggplot(data = swk, aes(x = year, y = broad_evg_clop_patch_density, color = district)) +  
  geom_line() +
  labs(title = 'Patch Density, Sarawak')

ggplot(data = sbh, aes(x = year, y = broad_evg_clop_patch_density, color = district)) +  
  geom_line() +
  labs(title = 'Patch Density, Sabah')

#Patch density
ggplot(data = swk, aes(x = year, y = broad_evg_clop_prop, color = district)) +  
  geom_line() +
  labs(title = 'Proportion Forest Cover, Sarawak')

ggplot(data = sbh, aes(x = year, y = broad_evg_clop_prop, color = district)) +  
  geom_line() +
  labs(title = 'Proportion Forest Cover, Sabah')
```

Agricultural cover across districts, Sarawak and Sabah
```{r}
#Total area
ggplot(data = swk, aes(x = year, y = agri_total_area, color = district)) +  
  geom_line() +
  labs(title = 'Total Agricultural Cover, Sarawak')

ggplot(data = sbh, aes(x = year, y = agri_total_area, color = district)) +  
  geom_line() +
  labs(title = 'Total Agricultural Cover, Sabah')

#Proportion
ggplot(data = swk, aes(x = year, y = agri_prop, color = district)) +  
  geom_line() +
  labs(title = 'Proportion Agricultural Cover, Sarawak')

ggplot(data = sbh, aes(x = year, y = agri_prop, color = district)) +  
  geom_line() +
  labs(title = 'Proportion Agricultural Cover, Sabah')
```

# Correlations

```{r}
res_sub <- rcorr(as.matrix(sbh_sub))

corrplot(res_sub$r, type = "lower", 
        p.mat = res_sub$P, sig.level = 0.05, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex = 0.5)

res_sub2 <- rcorr(as.matrix(swk_sub))

corrplot(res_sub2$r, type = "lower", 
        p.mat = res_sub2$P, sig.level = 0.05, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex=0.5)
```

Take into account year
```{r}
sbh_sub_time = sbh %>% 
  filter(year == 2000) %>% 
  select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "shrub_evg_prop", "shrub_evg_n_patches", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density","crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "shrub_evg_patch_density", "prec_mean"))

res_sub_time = rcorr(as.matrix(sbh_sub_time))

corrplot(res_sub_time$r, type = "lower", 
        p.mat = res_sub_time$P, sig.level = 0.05, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex = 0.5)  

swk_sub_time = swk %>% 
  filter(year == 2010) %>% 
  select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "shrub_evg_prop", "shrub_evg_n_patches", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density","crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "shrub_evg_patch_density", "prec_mean"))

res_sub_time2 = rcorr(as.matrix(swk_sub_time))

corrplot(res_sub_time2$r, type = "lower", 
        p.mat = res_sub_time2$P, sig.level = 0.05, insig = "blank", tl.col = "black", tl.cex = 0.5, cl.cex = 0.5)
```

# Models

Split data into training and testing sets
```{r}
train_nabawan = sbh %>% 
  filter(year %in% c(2000, 2001, 2002, 2003, 2005, 2006, 2008, 2009, 2010))

test_sbh = sbh %>% 
  filter(year %in% c(2011, 2012, 2013)) 

train_swk = swk %>% 
  filter(year %in% c(2010, 2011, 2012, 2013))

test_swk = swk %>% 
  filter(year %in% c(2014)) 
```

### Sabah

```{r}
smr_var_sbh = sbh %>% 
  group_by(district) %>% 
  summarise(variance = var(smr)) %>% 
  knitr::kable()

smr_var_sbh
```


spatial autocorrelation
```{r}
library(raster)
library(spdep)
malaysia = getData('GADM', country='MYS', level=2)
sabah = malaysia[malaysia$NAME_1 == 'Sabah',]

nb <- poly2nb(sabah, row.names = sabah$NAME_2)
names(nb) <- attr(nb, "region.id")

```

GAM
```{r}
#library(MASS)
set.seed(1)

sbh_forest.gam = gam(smr ~ te(broad_evg_clop_prop, year) +
                       #te(broad_evg_clop_n_patches, year) +
                       #te(broad_evg_clop_patch_density, year) +
                       s(year, k = 9) +
                       s(district, bs = 'mrf', xt = list(nb = nb))
                     ,
                     data = train_sbh, method = 'REML'
                     #, family = nb(theta = NULL, link = "log")
                     )

#plot(sbh_forest.gam)
#summary(sbh_forest.gam)
#gam.check(sbh_forest.gam)

sbh_agri.gam = gam(smr ~ te(agri_prop, year) +
                       #te(crop_rainfed_n_patches, year) +
                       #te(crop_rainfed_patch_density, year) +
                       #te(crop_irrigate_n_patches, year) +
                       #te(crop_irrigate_patch_density, year) +
                       #te(mos_crop_n_patches, year) +
                       #te(mos_crop_patch_density, year) +
                       #te(mos_natural_n_patches, year) +
                       #te(mos_natural_patch_density, year) +
                       s(year, k = 9) +
                       s(district, bs = 'mrf', xt = list(nb = nb))
                   ,
                     data = train_sbh, method = 'REML'
                   #, family = nb(theta = NULL, link = "log")
                   )

#plot(sbh_agri.gam)
#summary(sbh_agri.gam)
#gam.check(sbh_agri.gam)

sbh_water.gam = gam(smr ~ te(prec_mean, year) +
                      #te(tree_shrub_patch_density, year) +
                      #te(flood_fresh_n_patches, year) +
                      #te(flood_fresh_patch_density, year) +
                      #te(flood_fresh_prop, year) +
                      #te(flood_salt_n_patches, year) +
                      #te(flood_salt_patch_density, year) +
                      #te(flood_salt_prop, year) +
                      s(year, k = 9) +
                      s(district, bs = 'mrf', xt = list(nb = nb))
                    ,
                    data = train_sbh, method = 'REML'
                    #, family = nb(theta = NULL, link = "log")
                    )

#vis.gam(sbh_agri.gam, plot.type = 'contour')
#plot(sbh_water.gam)
#summary(sbh_water.gam)
#gam.check(sbh_water.gam)

sbh_urban.gam = gam(smr ~ te(urban_prop, year) +
                       #te(urban_n_patches, year) +
                       #te(urban_patch_density, year) +
                       s(year, k = 9) +
                       s(district, bs = 'mrf', xt = list(nb = nb))
                     ,
                     data = train_sbh, method = 'REML'
                     #, family = nb(theta = NULL, link = "log")
                     )

pred_forest = predict(sbh_forest.gam, newdata = test_sbh)
pred_agri = predict(sbh_agri.gam, newdata = test_sbh)
pred_water = predict(sbh_water.gam, newdata = test_sbh)
pred_urban = predict(sbh_urban.gam, newdata = test_sbh)
```

```{r}
mean(data.matrix(test_sbh[,57] - pred_forest)^2)
mean(data.matrix(test_sbh[,57] - pred_agri)^2)
mean(data.matrix(test_sbh[,57] - pred_water)^2)
mean(data.matrix(test_sbh[,57] - pred_urban)^2)
```

```{r}
pred_sbh = data.matrix((pred_forest + pred_agri + pred_water + pred_urban)/4)

mean(data.matrix(test_sbh[,57] - pred_sbh)^2)
```


```{r}
ctrl1 <- trainControl(method = "repeatedcv", number = 10, repeats = 5)

sbh_sub_tr = sbh %>% 
  filter(year %in% c(2000, 2001, 2002, 2003, 2005, 2006, 2008, 2009, 2010)) %>% 
  dplyr::select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "shrub_evg_prop", "shrub_evg_n_patches", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density", "crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "shrub_evg_patch_density", "prec_mean", "year", "district"))

sbh_sub_te = swk %>% 
  filter(year %in% c(2011, 2012, 2013)) %>% 
  dplyr::select(c("smr", "broad_evg_clop_prop", "crop_rainfed_prop", "herbaceous_prop", "shrub_prop", "urban_prop", "tree_shrub_prop", "mos_crop_prop", "mos_natural_prop", "sparse_veg_prop", "flood_fresh_prop", "flood_salt_prop", "water_prop", "crop_irrigate_prop", "shrub_evg_prop", "shrub_evg_n_patches", "broad_evg_clop_n_patches", "crop_rainfed_n_patches", "herbaceous_n_patches", "shrub_n_patches", "urban_n_patches", "tree_shrub_n_patches", "mos_crop_n_patches", "mos_natural_n_patches", "sparse_veg_n_patches", "flood_fresh_n_patches", "flood_salt_n_patches", "water_n_patches", "crop_irrigate_n_patches", "broad_evg_clop_patch_density", "crop_rainfed_patch_density", "herbaceous_patch_density", "shrub_patch_density", "urban_patch_density", "tree_shrub_patch_density", "mos_crop_patch_density", "mos_natural_patch_density", "sparse_veg_patch_density", "flood_fresh_patch_density", "flood_salt_patch_density", "water_patch_density", "crop_irrigate_patch_density", "shrub_evg_patch_density", "prec_mean", "year", "district")) %>% 


tr_sbh.x = model.matrix(smr~., sbh_sub_tr, na.action = NULL)[,-1]
tr_sbh.y = sbh_sub_tr$smr

te_sbh.x = model.matrix(smr~., sbh_sub_te)[,-1]
te_sbh.y = sbh_sub_te$smr
```

LASSO
```{r warning=FALSE, results='hide'}
set.seed(1)
lasso.fit = train(tr_sbh.x, tr_sbh.y, 
              method = "glmnet",
              tuneGrid = expand.grid(alpha = 1, 
                                     lambda = exp(seq(-10, 10, length = 200))),
              trControl = ctrl1)
plot(lasso.fit, xTrans = function(x) log(x))
lasso.fit$bestTune
coef(lasso.fit$finalModel,lasso.fit$bestTune$lambda)

best.lambda.lasso = lasso.fit$bestTune$lambda

lasso.pred = predict(lasso.fit$finalModel, s = best.lambda.lasso, newx = te_sbh.x)

mean((lasso.pred - te_sbh.y)^2)
```

Ridge
```{r}
set.seed(1)
ridge.fit = train(tr_sbh.x, tr_sbh.y, 
              method = "glmnet",
              tuneGrid = expand.grid(alpha = 0, 
                                     lambda = exp(seq(-10, 10, length = 200))),
              trControl = ctrl1)
plot(ridge.fit, xTrans = function(x) log(x))
ridge.fit$bestTune
coef(ridge.fit$finalModel,ridge.fit$bestTune$lambda)

best.lambda.ridge <- ridge.fit$bestTune$lambda

ridge.pred = predict(ridge.fit$finalModel, s = best.lambda.ridge, newx = te_sbh.x)

mean((ridge.pred - te_sbh.y)^2)
```

```{r}
set.seed(1)
enet.fit <- train(sbh.x, sbh.y,
                     method = "glmnet",
                     tuneGrid = expand.grid(alpha = seq(0, 1, length = 5), 
                                            lambda = exp(seq(-2, 4, length=50))),
                   # preProc = c("center", "scale"),
                     trControl = ctrl1)
enet.fit$bestTune

ggplot(enet.fit)
```

```{r}
resamp <- resamples(list(lasso = lasso.fit, ridge = ridge.fit, enet = enet.fit))
summary(resamp)

parallelplot(resamp, metric = "RMSE")
bwplot(resamp, metric = "RMSE")
```

graph
```{r}
ggplot()
```


### Sarawak


spatial autocorrelation
```{r}
library(raster)
library(spdep)
malaysia = getData('GADM', country='MYS', level=2)
sarawak = malaysia[malaysia$NAME_1 == 'Sarawak',]

nb2 <- poly2nb(sarawak, row.names = sarawak$NAME_2)
names(nb2) <- attr(nb2, "region.id")

```


GAM
```{r}
library(MASS)
set.seed(1)

swk_forest.gam = gam(smr ~ te(broad_evg_clop_prop, year, k = c(10, 3)) +
                       #te(broad_evg_clop_n_patches, year, k = c(10, 3)) +
                       #te(broad_evg_clop_patch_density, year, k = c(10, 3)) +
                       s(year, k = 3) 
                       #s(as.factor(district), bs = 'mrf', xt = list(nb = nb2))
                     ,
                     data = train_swk, method = 'REML'
                     #, family = nb(theta = NULL, link = "log")
                     )

#plot(swk_forest.gam)
#summary(swk_forest.gam)
#gam.check(swk_forest.gam)

swk_agri.gam = gam(smr ~ te(agri_prop, year, k = c(5, 3)) +
                       #te(crop_rainfed_n_patches, year, k = c(5, 3)) +
                       #te(crop_rainfed_patch_density, year, k = c(5, 3)) +
                       #te(crop_irrigate_n_patches, year, k = c(5, 3)) +
                       #te(crop_irrigate_patch_density, year, k = c(5, 3)) +
                       #te(mos_crop_n_patches, year, k = c(5, 3)) +
                       #te(mos_crop_patch_density, year, k = c(5, 3)) +
                       #te(mos_natural_n_patches, year, k = c(5, 3)) +
                       #te(mos_natural_patch_density, year, k = c(5, 3)) +
                       s(year, k = 3) 
                       #s(district, bs = 'mrf', xt = list(nb = nb))
                   ,
                     data = train_swk, method = 'REML'
                   #, family = nb(theta = NULL, link = "log")
                   )

#plot(swk_agri.gam)
#summary(swk_agri.gam)
#gam.check(swk_agri.gam)

swk_water.gam = gam(smr ~ te(prec_mean, year, k = c(5, 3)) +
                      #te(tree_shrub_patch_density, year, k = c(5, 3)) +
                      #te(flood_fresh_n_patches, year, k = c(5, 3)) +
                      #te(flood_fresh_patch_density, year, k = c(5, 3)) +
                      #te(flood_fresh_prop, year, k = c(5, 3)) +
                      #te(flood_salt_n_patches, year, k = c(5, 3)) +
                      #te(flood_salt_patch_density, year, k = c(5, 3)) +
                      #te(flood_salt_prop, year, k = c(5, 3)) +
                      s(year, k = 3) 
                      #s(district, bs = 'mrf', xt = list(nb = nb))
                    ,
                    data = train_swk, method = 'REML'
                    #, family = nb(theta = NULL, link = "log")
                    )

#plot(swk_water.gam)
#summary(swk_water.gam)
#gam.check(swk_water.gam)

swk_urban.gam = gam(smr ~ te(urban_prop, year, k = c(5, 3)) +
                       #te(urban_n_patches, year, k = c(5, 3)) +
                       #te(urban_patch_density, year, k = c(5, 3)) +
                       s(year, k = 3)
                       #s(district, bs = 'mrf', xt = list(nb = nb))
                     ,
                     data = train_swk, method = 'REML'
                     #, family = nb(theta = NULL, link = "log")
                     )

pred_forest2 = predict(swk_forest.gam, newdata = test_swk)
pred_agri2 = predict(swk_agri.gam, newdata = test_swk)
pred_water2 = predict(swk_water.gam, newdata = test_swk)
pred_urban2 = predict(swk_urban.gam, newdata = test_swk)
```

```{r}
mean(data.matrix(test_swk[,5] - pred_forest2)^2)
mean(data.matrix(test_swk[,5] - pred_agri2)^2)
mean(data.matrix(test_swk[,5] - pred_water2)^2)
mean(data.matrix(test_swk[,5] - pred_urban2)^2)
```

```{r}
pred_swk = data.matrix((pred_forest2 + pred_agri2 + pred_water2 + pred_urban2)/4)

mean(data.matrix(test_swk[,5] - pred_swk)^2)
```

Try "prophet" package
